---
title: "Brazil - Pesticide and Honey Production"
output: github_document
bibliography: refs.bib 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)


```

Honey bees have a reputation for working hard. We are working hard to cause bee losses. From 1966 to 1979, most bee losses were attributable to pesticides (@ricscu). Ignoring the impact of pesticides on bee lives, in recent years, Brazil introduced a staggering number of new pesticides. So let's look at data about honey production across Brazilian estates, and pesticides.

The data about honey production is from the Municipal Livestock Research collected by the Brazilian Institute of Geography and Statistics (IBGE), and data about approval of new pesticides is from the Brazilian Ministry of Agriculture and Livestock. Our analysis will target the converging timestamp in both datasets: 2000-2021.

```{r honey_data}


get_honey_data <- function(){
  
  url_honey_data <- "https://sidra.ibge.gov.br/geratabela?format=xlsx&name=tabela74.xlsx&terr=NS&rank=-&query=t/74/n1/all/n3/all/u/y/v/all/p/all/c80/0,2687/d/v1000215%202/l/v,p%2Bc80,t"
  
  folder <- "data"
  file <- file.path(folder, "honey_ppm_ibge.xlsx")
  data_processed <- file.path(folder, "honey_processed.csv")
  
  if(dir.exists(folder) == FALSE) dir.create(folder)
  
  if(file.exists(file) == FALSE) download.file(url_honey_data, file)
  
  if(file.exists(data_processed) == TRUE){
    data_final <- read_csv(data_processed)
  }else{
    production_value <-  readxl::read_excel("data/honey_ppm_ibge.xlsx", 
                                            sheet = "Valor da produção", skip = 4)
    
    production_kg <- readxl::read_excel("data/honey_ppm_ibge.xlsx", 
                                        skip = 4)
    # select cols about honey and estate id
    data <- list(production_value = production_value,
                 production_kg = production_kg) %>% 
      lapply(., function(df) df %>%  select(2, contains("Mel")))
    
    # New col names
    new_col_names <- list()
    new_col_names$value <- c("estate", paste("year", 1974:2021, sep = "_"))
    new_col_names$honey <- c("estate", paste("year", 1974:2021, sep = "_"))
    
    # correct col names
    data_adj <- map2(data, new_col_names, function(data, col_names){
      colnames(data) <- col_names
      
      data <- data %>% 
        pivot_longer(cols = c(everything(), -c("estate")),
                     names_to = "year")
      
      return(data)
    })
    
    data_adj$production_kg <- data_adj$production_kg %>% rename(honey = value)
    
    data_final <- data_adj %>%
      purrr::reduce(full_join) %>% 
      mutate(year = gsub("year_", "", year),
             # according to the data "-" is 0
             across(c(everything(), - estate), ~gsub("-", 0, .)),
             across(c(everything(), - estate), ~as.numeric(.)))
    
    write_csv(data_final, "data/honey_processed.csv")
    
  }
  
  return(data_final)
  
  

}

honey_data <- get_honey_data()


```

```{r pesticide_approval_data}



get_pesticide_approval_data <- function(){
  
  url_pesticide <- "https://www.gov.br/agricultura/pt-br/assuntos/insumos-agropecuarios/insumos-agricolas/agrotoxicos/RegistrosConcedidos20002023.xlsx"
  
  folder <- "data"
  file <- file.path(folder, "approved_pesticide.xlsx")
  data_processed <- file.path(folder, "pesticide_processed.csv")
  
  if(file.exists(file) == FALSE) download.file(url_pesticide, file)
  
  pest_data <- readxl::read_excel(file, 
                             sheet = "Tabela Geral 2020 e Outros", 
                             skip = 2)
  
  #This dataset is splited in several tables inside the same sheet. They are separated by blank rows. We will use this pattern to extract the interesting tables
  
  cols_until_2021 <- grep("2021", x = colnames(pest_data))
  
  pest_data_correct_cols <- pest_data %>% select(1:cols_until_2021, -contains("..."))
  
  seq_index <- function(vec){
  #https://stackoverflow.com/questions/29184297/finding-the-start-and-stop-indices-in-sequence-in-r
    start<- vec[!(vec-1) %in% vec]
    stop <- vec[!(vec+1) %in% vec]
    cbind(start, stop) %>% 
      as_tibble()
  }
  
  NAs_rows <-  which(!is.na(pest_data_correct_cols$"2000.0"))

  tabls_ind <- seq_index(NAs_rows)

  #ANVISA assessment
  toxic_risk <- pest_data_correct_cols %>% 
    slice(tabls_ind$start[1]:tabls_ind$stop[1])

  #IBAMA assessment
  environmental_risk <- pest_data_correct_cols %>%
    slice(tabls_ind$start[2]:tabls_ind$stop[2])

  list(toxic_risk, environmental_risk)
  
  
}

if(file.exists("data/ambiental_hazard_approv.csv") == TRUE){
 ambiental_hazard_approv <- read_csv("data/ambiental_hazard_approv.csv") 
}else{
  pesticide_approval <- get_pesticide_approval_data()
  
  ambiental_hazard_approv <- pesticide_approval[[2]] %>% 
  rename(hazard_classf = 1) %>% 
  slice(-1) %>% 
  pivot_longer(cols = -hazard_classf, 
               values_transform = as.character,
               names_to = "year",
               values_to = "n_apprvd") %>% 
  mutate(across(-hazard_classf, ~as.numeric(.)))
  
  write.csv(ambiental_hazard_approv, "data/ambiental_hazard_approv.csv", row.names = FALSE)

}





```

```{r pesticide_purchase}

get_pesticide_purchase_data <- function(){
  
  folder <- "data"
  
  files_download <- lapply(2009:2021, function(year, folder){
    
    url <- sprintf("https://www.gov.br/ibama/pt-br/assuntos/quimicos-e-biologicos/agrotoxicos/arquivos/qualidadeambiental/relatorios/%i/", year)
    
    html <- rvest::read_html(url) %>% 
      rvest::html_elements("a") %>% 
      rvest::html_text()
    
    file_get <- html[grepl("classe_periculosidade.*xls|classe_.*ambiental.*xls", html)][1]
    
    url_get <- paste0(url, file_get)
    
    file_name <- sprintf("Vendas_por_classe_de_uso_%i.xls", year)
    file_path <- file.path(folder, file_name)
    
    if(file.exists(file_path) == FALSE) download.file(url_get, file_path)
    
    data <- readxl::read_excel(file_path, range = "A4:H39")
    
    data <- data %>%
      select(1, contains("CLASSE")) %>%
      slice(-1) %>%
      rename(location = 1) %>%
      pivot_longer(- location,
                   names_to = "class",
                   values_to = "purchase_ton") %>%
      mutate(year = year)
    
    return(data)
    
    
  }, folder = folder)

  
  return(files_download)
  
    

}

pesticide_purchase <- get_pesticide_purchase_data()%>%
  plyr::ldply("tibble")

```


# References

<div id="refs"></div>


