---
title: "Brazil - Pesticide and Honey Production"
author: "Luiz Felipe Martucci"
output: github_document
bibliography: refs.bib 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include = FALSE}
req_libs <- c("tidyverse", 
              "readxl",
              "plyr",
              "purrr",
              "geojsonio",
              "geogrid",
              "rgdal",
              "ggtext",
              "rgeos",
              "geojsonio",
              "broom",
              "forcats",
              "readr"
)

b <- (function(libs){  
  sapply(libs, function(x){
    if(!x %in% installed.packages()) install.packages(x, dependencies = T)
  })
  
  lib_version <- sapply(libs, function(x){
    lib_version <- getNamespaceVersion(x)
    string_lib <- paste0(x, " installed version: ", lib_version)
    print(string_lib)
    
  })
  
  lib_version <- sort(lib_version)
  
  today_date <- date()
  r_version <- R.Version()$version.string
  
  requirements <- c(today_date, 
                    osVersion,
                    r_version,
                    lib_version)
  
  writeLines(requirements, "requirements.txt")

    
return(requirements)
  
  
})(req_libs)
  
library(tidyverse)


```


Honey bees have a reputation for working hard. They not only produce honey but pollinate flowering plants. Together with other pollinators, they are responsible for more than 87% of flowering plant species' reproduction (@ostiguy). However, we are working hard to cause bee losses. From 1966 to 1979, most bee losses were attributable to pesticides (@ricscu), especially pesticides applied to agricultural or horticultural landscapes (@ostiguy). Ignoring their impact on bee lives, Brazil has been introducing a staggering amount of new pesticides year after year. So let's look at data about honey production across Brazilian states and the number of new pesticides approved.


The data about honey production is from the Municipal Livestock Research collected by the Brazilian Institute of Geography and Statistics (IBGE), and data about approval of new pesticides is from the Brazilian Ministry of Agriculture and Livestock. Our analysis will target the converging timestamp in both datasets: 2000-2021.

```{r honey_data}


get_honey_data <- function(){
  
  url_honey_data <- "https://sidra.ibge.gov.br/geratabela?format=xlsx&name=tabela74.xlsx&terr=NS&rank=-&query=t/74/n1/all/n3/all/u/y/v/all/p/all/c80/0,2687/d/v1000215%202/l/v,p%2Bc80,t"
  
  folder <- "data"
  file <- file.path(folder, "honey_ppm_ibge.xlsx")
  data_processed <- file.path(folder, "honey_processed.csv")
  
  if(dir.exists(folder) == FALSE) dir.create(folder)
  
  if(file.exists(file) == FALSE) download.file(url_honey_data, file)
  
  if(file.exists(data_processed) == TRUE){
    data_final <- read_csv(data_processed)
  }else{
    production_value <-  readxl::read_excel("data/honey_ppm_ibge.xlsx", 
                                            sheet = "Valor da produção", skip = 4)
    
    production_kg <- readxl::read_excel("data/honey_ppm_ibge.xlsx", 
                                        skip = 4)
    # select cols about honey and state id
    data <- list(production_value = production_value,
                 production_kg = production_kg) %>% 
      lapply(., function(df) df %>%  select(2, contains("Mel")))
    
    # New col names
    new_col_names <- list()
    new_col_names$value <- c("state", paste("year", 1974:2021, sep = "_"))
    new_col_names$honey <- c("state", paste("year", 1974:2021, sep = "_"))
    
    # correct col names
    data_adj <- map2(data, new_col_names, function(data, col_names){
      colnames(data) <- col_names
      
      data <- data %>% 
        pivot_longer(cols = c(everything(), -c("state")),
                     names_to = "year")
      
      return(data)
    })
    
    data_adj$production_kg <- data_adj$production_kg %>% rename(honey = value)
    
    data_final <- data_adj %>%
      purrr::reduce(full_join) %>% 
      mutate(year = gsub("year_", "", year),
             # according to the data "-" is 0
             across(c(everything(), - state), ~gsub("-", 0, .)),
             across(c(everything(), - state), ~as.numeric(.)))
    
    write_csv(data_final, "data/honey_processed.csv")
    
  }
  
  return(data_final)
  
  

}

honey_data <- get_honey_data()

honey_data <- honey_data %>% 
  filter(year %in% 2000:max(year) & 
           !state %in% c("Guanabara", NA)) %>% 
  mutate(profit_by_production = value / honey)


```

```{r pesticide_approval_data}



get_pesticide_approval_data <- function(){
  
  url_pesticide <- "https://www.gov.br/agricultura/pt-br/assuntos/insumos-agropecuarios/insumos-agricolas/agrotoxicos/RegistrosConcedidos20002023.xlsx"
  
  folder <- "data"
  file <- file.path(folder, "approved_pesticide.xlsx")
  data_processed <- file.path(folder, "pesticide_processed.csv")
  
  if(file.exists(file) == FALSE) download.file(url_pesticide, file)
  
  pest_data <- readxl::read_excel(file, 
                             sheet = "Tabela Geral 2020 e Outros", 
                             skip = 2)
  
  #This dataset is splited in several tables inside the same sheet. They are separated by blank rows. We will use this pattern to extract the interesting tables
  
  cols_until_2021 <- grep("2021", x = colnames(pest_data))
  
  pest_data_correct_cols <- pest_data %>% select(1:cols_until_2021, -contains("..."))
  
  seq_index <- function(vec){
  #https://stackoverflow.com/questions/29184297/finding-the-start-and-stop-indices-in-sequence-in-r
    start<- vec[!(vec-1) %in% vec]
    stop <- vec[!(vec+1) %in% vec]
    cbind(start, stop) %>% 
      as_tibble()
  }
  
  NAs_rows <-  which(!is.na(pest_data_correct_cols$"2000.0"))

  tabls_ind <- seq_index(NAs_rows)

  #ANVISA assessment
  toxic_risk <- pest_data_correct_cols %>% 
    slice(tabls_ind$start[1]:tabls_ind$stop[1])

  #IBAMA assessment
  environmental_risk <- pest_data_correct_cols %>%
    slice(tabls_ind$start[2]:tabls_ind$stop[2])

  list(toxic_risk, environmental_risk)
  
  
}

if(file.exists("data/ambiental_hazard_approv.csv") == TRUE){
 ambiental_hazard_approv <- read_csv("data/ambiental_hazard_approv.csv") 
}else{
  pesticide_approval <- get_pesticide_approval_data()
  
  ambiental_hazard_approv <- pesticide_approval[[2]] %>% 
  rename(hazard_classf = 1) %>% 
  slice(-1) %>% 
  pivot_longer(cols = -hazard_classf, 
               values_transform = as.character,
               names_to = "year",
               values_to = "n_apprvd") %>% 
  mutate(across(-hazard_classf, ~as.numeric(.)))
  
  write.csv(ambiental_hazard_approv, "data/ambiental_hazard_approv.csv", row.names = FALSE)

}





```

```{r pesticide_purchase}

get_pesticide_purchase_data <- function(){
  
  folder <- "data"
  
  files_download <- lapply(2009:2021, function(year, folder){
    
    url <- sprintf("https://www.gov.br/ibama/pt-br/assuntos/quimicos-e-biologicos/agrotoxicos/arquivos/qualidadeambiental/relatorios/%i/", year)
    
    html <- rvest::read_html(url) %>% 
      rvest::html_elements("a") %>% 
      rvest::html_text()
    
    file_get <- html[grepl("classe_periculosidade.*xls|classe_.*ambiental.*xls", html)][1]
    
    url_get <- paste0(url, file_get)
    
    file_name <- sprintf("Vendas_por_classe_de_uso_%i.xls", year)
    file_path <- file.path(folder, file_name)
    
    if(file.exists(file_path) == FALSE) download.file(url_get, file_path)
    
    data <- readxl::read_excel(file_path, range = "A4:H39")
    
    data <- data %>%
      select(1, contains("CLASSE")) %>%
      slice(-1) %>%
      rename(location = 1) %>%
      pivot_longer(- location,
                   names_to = "class",
                   values_to = "purchase_ton") %>%
      mutate(year = year)
    
    return(data)
    
    
  }, folder = folder)

  
  return(files_download)
  
    

}

if(file.exists("data/pesticide_purchases.csv") == TRUE){
  pesticide_purchase <- read_csv("data/pesticide_purchases.csv")
}else{
  pesticide_purchase <- get_pesticide_purchase_data()%>%
  plyr::ldply("tibble")
  
  write.csv(pesticide_purchase, "data/pesticide_purchases.csv", row.names = FALSE)
}


```


```{r state_info}

if(file.exists("data/states_info.csv")== TRUE){
  states_info <- read_csv("data/states_info.csv")
}else{
  get_states_info <- function(){
    
   url <-  "http://blog.mds.gov.br/redesuas/wp-content/uploads/2018/06/Lista_Estados_Brasil_Versao_CSV.csv"
   
   download.file(url, "data/states_info.csv")
   
   read_csv("data/states_info.csv")
   
  }

states_info <- get_states_info()
  
}



```



```{r map_data}


if(file.exists("data/tidy_brazil_map.csv") == FALSE){
  make_brazil_map <- function(){
    folder <- "data"
    brazil_raw_map <- file.path(folder, "brazil_map.json") 
    
    if(file.exists(brazil_raw_map) == FALSE){
      download.file("https://gist.githubusercontent.com/ruliana/1ccaaab05ea113b0dff3b22be3b4d637/raw/196c0332d38cb935cfca227d28f7cecfa70b412e/br-states.json",
                    brazil_raw_map)
    }
    
    og_shape <- geojsonio::geojson_read(brazil_raw_map, what="sp")
    
    og_shape@data <- og_shape@data %>% rename(state = nome)
    
    
    #The map comes without id on @data
    #Corrects id on @data and calculate center position for each UF
    centers <-  data.frame(og_shape %>% rgeos::gCentroid(byid = T)) %>% 
  mutate(id = row_number() %>% as.character(),
         UF = c("AC", "AL", "AM", "AP", "BA", "CE",
         "DF", "ES", "GO", "MA", "MG", "MS",
         "MT", "PA", "PB", "PE", "PI", "PR", 
         "RJ", "RN", "RO", "RR", "RS", "SC",
         "SE", "SP", "TO"))
    
    tidy_map <- og_shape %>% 
      broom::tidy() %>% 
      left_join(centers)
    
    write.csv(tidy_map, "data/tidy_brazil_map.csv", row.names = FALSE)
    
    return(tidy_map)
    
  }
  
  tidy_map <- make_brazil_map()
  
}else{
  tidy_map <- read_csv("data/tidy_brazil_map.csv")
  
}










```

# References

<div id="refs"></div>


